Identifying the Tissue of Origin of Extracellular Vesicles Using RNA Expression Signatures

### 2015-09-08 poster/poster.Rnw
start working on the introduction section by adapting the abstract

  \block{Abstract}{ Extracellular vesicles (EVs) are small (40-1000nm)
    compartments that can be found in blood, saliva, and urine. EVs
    originate from the plasma membrane or multivesicular endosome of
    many cell types and contain proteins, lipids, mRNA, miRNA, and
    many other non-coding RNAs. Because EVs contain mRNA from the cell
    of origin and can be obtained without invasive procedures, they
    are a promising remote sensor of the transcriptome of tissues
    which would otherwise be inaccessible. One requirement for
    remotely sensing the transcriptome is identifying the source
    tissue of a specific EV. To identify the source tissue of a
    specific EV, we have identified tissue-specific miRNA and mRNA
    markers for multiple organs and tissues from existing
    publicly-available RNAseq transcriptomes by calculating a tissue
    specificity index. Using machine learning techniques we trained a
    multiclass support vector machine to classify the tissue of
    origin. Our approach identified MIR1914, MIR126, SNORD12, and
    SNORD5 as uterus specific. CGB, LGALS13, and members of the PSG
    family were identified as placenta specific. The accuracy of the
    classifier was verified using RNAseq transcriptome data not in the
    training set. We show the applicability of the classifier to
    simulated EVs by down sampling reads to $10^5$ reads and biasing
    the reads to account for preferential association of RNA motifs
    with EV packaging machinery. When coupled with single-EV
    sequencing, our classifier enables clinicians to monitor
    disease-specific markers over the course of treatment without
    biopsies, increasing the ability of clinicians to adapt treatment
    modalities to the individual patient. }

>>>>

  \block{Introduction}{\acp{EV} are small (40-1000nm) compartments
    that can be found in blood, saliva, and urine. \acp{EV} originate
    from the plasma membrane or multivesicular endosome of many cell
    types and contain proteins, lipids, mRNA, miRNA, and many other
    non-coding RNAs. Because \acp{EV} contain mRNA from the cell of
    origin and can be obtained without invasive procedures, they are a
    promising remote sensor of the transcriptome of tissues which
    would otherwise be inaccessible.

    One requirement for remotely sensing the transcriptome is
    identifying the source tissue of a specific \ac{EV}. To identify
    the source tissue of a specific \ac{EV}, we have identified
    tissue-specific miRNA and mRNA markers for multiple organs and
    tissues from existing publicly-available RNAseq transcriptomes by
    calculating a tissue specificity index. Using machine learning
    techniques we trained a multiclass support vector machine to
    classify the tissue of origin.

%     Our approach identified MIR1914, MIR126, SNORD12, and SNORD5 as
%     uterus specific. CGB, LGALS13, and members of the PSG family were
%     identified as placenta specific. The accuracy of the classifier
%     was verified using RNAseq transcriptome data not in the training
%     set. We show the applicability of the classifier to simulated EVs
%     by down sampling reads to $10^5$ reads and biasing the reads to
%     account for preferential association of RNA motifs with EV
%     packaging machinery. When coupled with single-EV sequencing, our
%     classifier enables clinicians to monitor disease-specific markers
%     over the course of treatment without biopsies, increasing the
%     ability of clinicians to adapt treatment modalities to the
%     individual patient. 
  }
